---
title: Machine learning modelling of nutrition indicators in children under five years
  in Ghana- R version
output: html_document
date: '2022-06-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
pacman::p_load(
        tidyverse, # Data wrangling and visualisation
        skimr, # Skim dataset
        rio, # Import dataset
        here, # Directory management
        finalfit, # Model parameters and column labelling
        ggstatsplot, # Statistical plots
        GGally, # Scatterplot matrix
        caret, # Machine learning
        gtsummary, # Summary tables
        janitor, # Data cleaning
        parameters, # Model parameters
        naniar # Handling missing values
)
```

```{r import-dataset}
df <- import(here("data","GHKR72FL.SAV")) # Import dataset from the data folder
```



## Variable selection

- Divide the z-score variables by 100

#### Outcome variables

- WFH z-scores (HW72); 9996: height out of plausible limit, 9997: Age in days out of plausible limit, 9998: Flagged cases
- HFA z-scores (HW70); 9996: height out of plausible limit, 9997: Age in days out of plausible limit, 9998: Flagged cases
- WFA z-scores (HW71); 9996: height out of plausible limit, 9997: Age in days out of plausible limit, 9998: Flagged cases

#### Features

- Type of place of residence (V102): 1 for Urban, 2 for Rural
- child age in months (HW1)
- Birth order number (B0RD)
- Gender of the child (B4): 1 for Male, 2 for female
- Birth weight in kg (M19): 9996: No weight; 9998: Don't know; Divide by 1000
- Number of children 5 and under in household (V137)
- Highest education level (v106): 0 for No education, 1 for Primary, 2 for Secondary, 3 for Higher
- Wealth index (V190): 1 for Poorest, 2 for Poorer, 3 for Middle, 4 for Richer, 5 for Richest
- Had diarrhea recently (H11): 0 for No, 1 for Yes, last 24 hours, 2 for Yes, last two weeks, 8 for don't know (Join the values for yes into one)
- Type of toilet facility ();
- Weight in kg (HW2): 9994 for present, 9995 for Refused, 9996 for other; Divide weight by 10.
- Region (V024): 1 for Western, 2 for Central, 3 for Greater Accra, 4 for Volta, 5 for Eastern, 5 for Ashanti, 7 for Brong Ahafo, 8 for Nothern, 9 for Upper East, 10 for Upper West
- Source of drinking water (V113): 10 for Piped water, 20 for Tube well water, 32 for Unprotected well, 43 for River, 71 for Bottled water, 11 for Piped into dwelling, 21 for Tube well or borehole, 40 for Surface Water, 51 for rain water, 72 for Sachet water, 12 for Piped to yard/plot, 30 for Dug well, 41 for Protected spring, 61 for Tanker trunk, 96 for other, 13 for public tap/standpipe, 31 for protected well, 42 for unprotected spring, 62 for cart with small tank, and 97 for not a dejure resident   ; Recode to improved and non-improved sources of water




```{r column-selection}
# Use select() function from tidyverse to select columns of interest //TODO: Child's height inclusion, z-score flags
child <- df%>%
        select(HW72, #WFH z-scores
               HW70, # HFA z-scores
               HW71, # WFA z-scores
               V102, # Type of place of residence
               HW1, # Child age in months
               BORD, # Birth order number
               B4, # Gender of child
               M19, # Birth weight of child
               V137, # Number of children 5 and under
               V106, # Highest education level
               V190, # Wealth index
               H11, # Had diarrhoea recently
               HW2, # Weight of child
               V024, # Region
               V113) # Source of drinking water
```



```{r data-shape}
dim <- dim(child) # Dimensions of dataset
```

```{r glimpse-dataset}
ff_glimpse(child) #Overview of dataset including missing data
```



```{r drop-unusual-outcome}
# Dropped children with unusual nutritional indicators
child <- child%>%
        replace_with_na(
                replace = list(
                        HW72 = c(9996, 9997, 9998),
                        HW70 = c(9996, 9997, 9998),
                        HW71 = c(9996, 9997, 9998)
                )
        )
```


```{r glimpse-after-outcome-drop}
ff_glimpse(child)
```



```{r drop-missing}
child_final <- na.omit(child) # Drop missing observations
dim_new_df <- dim(child_final) # Assess dimensions of the dataset
dim_new_df
```

```{r replace-birthweight-not-weighed-dont-know}
# Replace unusual birthweight values
child_final <- child_final%>%
        replace_with_na(
                replace = list(M19 = c(9996, 9998)
                        
                )
        )
```


```{r replace-not-de-jeru-with-missing}
# Replace not de jeru in source of drinking water with missing
child_final <- child_final%>%
        replace_with_na(
                replace = list(V113 = 97
                        
                )
        )
```


```{r drop-missing-after-birthweight-replace}
child_final <- na.omit(child_final)
```


```{r glimpse-after-bwt-na}
ff_glimpse(child_final)
```

## Assessment of data

- There are `r dim[1]` observations in the dataset. The selected columns were `r dim[2]`.
- Data column names are not descriptive.
- Weight for height z-scores (HW72) has 3146 missing observations representing 53.5% of observations.
- Height for age z-scores (HW70) has 3146 missing observations representing 53.5% of observations.
- Weight for age z-scores (HW71) has 3146 missing observations representing 53.5% of observations.
- Child's age in month (HW1) has 3102 missing observations representing 52.7% of observations.
- Birth weight (M19) has 3 missing values representing 0.1% of observations.
- Had diarrheoa (H11) has 291 missing values representing 4.9% of observations. 
- Child's weight in kg has 3102 missing observations representing 52.7% of observations.
- After droping misssing cases, number ended up as `r dim_new_df[1]` 



```{r skim-dataset}
# Use skim() from skimr
# Note assessment of data
```



## Cleaning or recoding

- Remove missing values: Completed
- Change column names into descriptive names: Completed
- Divide z-score values (HW72, HW70, HW71) by 100: Completed
- Divide birthweight(M19) values by 1000 to convert into kg.
- Convert type of residence (V102), sex (B4), Highest educational level (V106), wealh index (V190), Had diarrhoea recently (H11), region (V024), source of drinking water (V113)  into factor type.
- Remove flaged z-score values.
- Mutate z-scores values into malnourished defined by z-score < -2 and all others as normal
- Remove unusual values from (HW72, HW70, HW71: 9996, 9997, 9998), (M19: 9996, 9998).



- WFH z-scores (HW72)
- HFA z-scores (HW70)
- WFA z-scores (HW71)
- Type of place of residence (V102): 1 for Urban, 2 for Rural
- child age in months (HW1)
- Birth order number (B0RD)
- Gender of the child (B4): 1 for Male, 2 for female
- Birth weight in kg (M19): 9996: No weight; 9998: Don't know; Divide by 1000
- Number of children 5 and under in household (V137)
- Highest education level (v106): 0 for No education, 1 for Primary, 2 for Secondary, 3 for Higher
- Wealth index (V190): 1 for Poorest, 2 for Poorer, 3 for Middle, 4 for Richer, 5 for Richest
- Had diarrhea recently (H11): 0 for No, 1 for Yes, last 24 hours, 2 for Yes, last two weeks, 8 for don't know (Join the values for yes into one)
- Type of toilet facility ();
- Weight in kg (HW2): 9994 for present, 9995 for Refused, 9996 for other; Divide weight by 10.
- Region (V024): 1 for Western, 2 for Central, 3 for Greater Accra, 4 for Volta, 5 for Eastern, 5 for Ashanti, 7 for Brong Ahafo, 8 for Nothern, 9 for Upper East, 10 for Upper West
- Source of drinking water (V113): 10 for Piped water, 20 for Tube well water, 32 for Unprotected well, 43 for River, 71 for Bottled water, 11 for Piped into dwelling, 21 for Tube well or borehole, 40 for Surface Water, 51 for rain water, 72 for Sachet water, 12 for Piped to yard/plot, 30 for Dug well, 41 for Protected spring, 61 for Tanker trunk, 96 for other, 13 for public tap/standpipe, 31 for protected well, 42 for unprotected spring, 62 for cart with small tank, and 97 for not a dejure resident   ; Recode to improved and non-improved sources of water
 

```{r rename-columns}
# Clean and preprocess data using janitor and tidyverse
# Rename column names using rename()
child_final <- child_final%>%
        rename(
                wfh = HW72,
                hfa = HW70,
                wfa = HW71,
                residence = V102,
                age = HW1,
                birth_order = BORD,
                gender = B4,
                bwt = M19,
                num_child = V137,
                educ_mum = V106,
                wealth_index = V190,
                diarrhea = H11,
                wt = HW2,
                region = V024,
                source_water = V113
                )
```


```{r division}
child_final<- child_final%>%
        mutate(
                # Divide nutritional indicators by 100
                wfh = wfh/100,
                hfa = hfa/100,
                wfa = wfa/100,
                # Divide birthweight
                bwt = bwt/1000)
```



```{r recoding}
# Recode variables
child_final <- child_final%>%
        mutate(
                # Categorical variables
                residence = factor(residence)%>%
                        fct_recode("Urban" = "1",
                                   "Rural" = "2")%>%
                        ff_label("Type of place of residence"),
                gender = factor(gender)%>%
                        fct_recode(
                                "Male" = "1",
                                "Female" = "2"
                        )%>%
                        ff_label("Gender of child"),
                educ_mum = factor(educ_mum)%>%
                        fct_recode(
                                "No education" = "0",
                                "Primary" = "1",
                                "Secondary" = "2",
                                "Higher" = "3"
                        )%>%
                        fct_relevel(
                                c(
                                        "No education",
                                        "Primary",
                                        "Secondary",
                                        "Higher"
                                )
                        )%>%
                        ff_label("Mother's educational level"),
                wealth_index = factor(wealth_index)%>%
                        fct_recode(
                                "Poorest" = "1",
                                "Poorer" = "2",
                                "Middle" = "3",
                                "Richer" = "4",
                                "Richest" = "5"
                        )%>%
                        fct_relevel(
                                c("Poorest",
                                  "Poorer",
                                  "Middle",
                                  "Richer",
                                  "Richest")
                        )%>%
                        ff_label("Household wealth index"),
                region = factor(region)%>%
                        fct_recode(
                                "Western" = "1",
                                "Central" ="2",
                                "Greater Accra" = "3",
                                "Volta" = "4",
                                "Eastern" = "5",
                                "Ashanti" = "6",
                                "Brong-Ahafo" = "7",
                                "Nothern" = "8",
                                "Upper East" = "9",
                                "Upper West" = "10"
                        )%>%
                        ff_label("Region of residence"),
                diarrhea = factor(diarrhea)%>%
                        fct_recode(
                                "No" = "0",
                                "Yes" = "1",
                                "Yes" = "2",
                                "Don't know" = "8"
                        )%>%
                        ff_label("Child has diarrhoea"),
                source_water = factor(source_water)%>%
                        fct_recode(
                                "Improved" = "10", # Pipe water
                                "Improved" = "20", # Tube well water
                                "Improved" = "71", # Bottle water
                                "Improved" = "11", # Piped into dwelling
                                "Improved" = "21", # Tube well/borehole
                                "Improved" = "72", # Sachet water
                                "Improved" = "12", # Piped into yard or plot
                                "Improved" = "30", # Dugged well (open/protected)
                                "Improved" = "31", # Protected well
                                "Improved" = "41", # Protected spring
                                "Improved" = "13", # Public tap/stand pipe
                                "Improved" = "51", # Rainwater
                                "Unimproved" = "32", # Unprotected well
                                "Unimproved" = "40", # Surface water
                                "Unimproved" = "42", # Unprotected spring
                                "Unimproved" = "43", # River/dam/pond/spring/irrigation/canal
                                "Unimproved" = "61", # Tanker truck
                                "Unimproved" = "62", # Cart with small tank
                                "Other" = "96"
                         )%>%
                        ff_label("Source of drinking water"),
                wfh_cat = ifelse(
                        wfh < -2, "Wasting",
                        "Normal"
                )%>%
                        ff_label("Wasting"),
                hfa_cat = ifelse(
                        hfa < -2, "Stunting",
                        "Normal"
                )%>%
                        ff_label("Stunting"),
                wfa_cat = ifelse(
                        wfa < -2, "Underweight",
                        "Normal"
                )%>%
                        ff_label("Underweight")
        )
```


```{r dummy-variables}
child_final<- child_final%>%
        mutate(
                wfh_dummy = wfh_cat%>%
                        fct_recode(
                                "0" = "Normal",
                                "1" = "Wasting"
                        ),
                
                hfa_dummy = hfa_cat%>%
                        fct_recode(
                                "0" = "Normal",
                                "1" = "Stunting"
                        ),
                
                wfa_dummy = wfa_cat%>%
                        fct_recode(
                                "0" = "Normal",
                                "1" = "Underweight"
                        )
        )
```


```{r glimpse-final-dataset}
ff_glimpse(child_final)
```

```{r export-clean-data}
export(child_final, here("data","clean_data.csv")) #Export clean dataset
```


## Exploratory data analysis


```{r statistical-visualisation}
# Stats visualisations using ggstatsplot
# Note observations under exploratory section
```


```{r scatterplot-matrix}
# Create a scatterplot matrix using ggpairs from GGally
# Note observations under exploratory section
```

## Introduction

Provide detailed information about the dataset

### Machine learning workflow

1.  Access and load the data: Completed
2.  Preprocess the data
3.  Derive features using the preprocessed data
4.  Train models using the features derived in step 3
5.  Iterate to find the best model
6.  Integrate the best-trained model into a production system

## Explanatory data analysis



## Baseline model


## Iterate


## Evaluate model
